---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data = read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
sum_data_per_day = aggregate(steps ~ date, data=data, sum, na.action=na.omit)
summary(sum_data_per_day)
hist(sum_data_per_day[,"steps"], breaks=20)
```

As can be seen by the result above the average steps/day is 10766.
The median is 10765.


## What is the average daily activity pattern?
```{r}
mean_data_per_interval = aggregate(steps ~ interval, data=data, mean, na.action=na.omit)
plot(mean_data_per_interval$interval, mean_data_per_interval$steps, type = "l")
```

```{r}
mean_data_per_interval[which.max( mean_data_per_interval[,"steps"] ),"interval"]
```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is the 835.


## Imputing missing values

The total number of missing values in the dataset (i.e. the total number of steps with NAs) is 2304:

```{r}
sum(is.na(data$steps))
sum(is.na(data$date))
sum(is.na(data$interval))
```

Let's fill NAs with global 5min interval mean, adjusting to the day's mean if not NA:

```{r}
typical_day_steps = mean(sum_data_per_day$steps)
sum_data_per_day_w_NA = aggregate(steps ~ date, data=data, sum, na.action=na.pass)
sum_data_relative_per_day_w_NA = sum_data_per_day_w_NA$steps/typical_day_steps

data_na_rows = which(is.na(data), arr.ind=TRUE)[,1]

filled_data = data
filled_data[data_na_rows, "steps"] = mean_data_per_interval[ mean_data_per_interval$interval == data[data_na_rows, "interval"], "steps"]
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


## Are there differences in activity patterns between weekdays and weekends?
